\emph{On Logics Extended with Embedding-Closed Quantifiers}

### 2020-11-21 paper.tex
Replace the homomorphism-closed quantifier section with a remark

\subsection{Homomorphism-closed quantifiers}

In addition to being closed under embeddings, we can think of various other
interesting closure conditions for quantifiers, like being closed under
homomorphisms for instance.
The purpose of this subsection is to show that any embedding-closed property can
be defined in a logic with homomorphism-closed quantifiers only.
Thus, requiring quantifiers to be closed under homomorphisms is not essentially
more restrictive than the requirement that they are closed under embeddings.
We denote by $\mathcal{Q}_{\hom}$ the class of all homomorphism-closed quantifiers.

\begin{theorem}
Let $\mathcal{L} = \mathcal{L}_{\kappa\omega}$ for some infinite cardinal
$\kappa$ or $\mathcal{L} = \mathcal{L}_{\infty\omega}^{\omega}$.
Then $\mathcal{L}(\mathcal{Q}_{\emb}) \equiv \mathcal{L}(\mathcal{Q}_{\hom})$.
\end{theorem}
\begin{proof}
Let $\tau$ be a relational vocabulary and $K$ an embedding-closed class of
$\tau$-structures.
Let $\tau' = \tau \cup \{R_* \colon R \in \tau \} \cup \{N\}$ with
$\ar(R_*)= \ar(R)$ for all $R \in\tau$ and $\ar(N) = 2$.
Let $F \colon \Str[\tau] \rightarrow \Str[\tau']$ be the function such that for
each $\mathfrak{A} \in \Str[\tau]$ the universe of
$F(\mathfrak{A})$ is $A$, $R^{F(\mathfrak{A})} = R^{\mathfrak{A}}$ and
$R_*^{F(\mathfrak{A})} = (\neg R)^{\mathfrak{A}}$ for all $R \in \tau$,
and $N^{F(\mathfrak{A})} = (x\ne y)^{\mathfrak{A}}$.
Now let
\[
	K' = \{\mathfrak{A}\in\tau' \colon \text{ there is } \mathfrak{B} \in K 
	\text{ and a homomorphism }f \colon F(\mathfrak{B})\rightarrow \mathfrak{A} \}.
\]
Then $K'$ is homomorphism-closed since the composition of two homomorphisms is
itself a homomorphism.
Let $\psi$ be the following sentence of $\mathcal{L}(\mathcal{Q}_{\hom})$:
\[
	\psi := Q_{K'}((\overline{x}_RR)_{R\in\tau},
	(\overline{y}_R\neg R)_{R\in\tau},uv(u\ne v)).
\]
We claim that $\psi$ defines the class $K$. To show this, suppose first that
$\mathfrak{A} \in \Str[\tau]$ is such that $\mathfrak{A}\vDash \psi$.
Then by the definition of $K'$ there is $\mathfrak{B} \in K$ and a homomorphism
\[
	f \colon F(\mathfrak{B}) \rightarrow 
	(A,(R^{\mathfrak{A}})_{R\in\tau},(\neg R^{\mathfrak{A}})_{R\in\tau},
	(x\ne y)^{\mathfrak{A}})
\]
which is in fact an embedding of $\mathfrak{B}$ into $\mathfrak{A}$ so,
since $K$ is embedding-closed, we have $\mathfrak{A} \in K$.
For the other direction, if $\mathfrak{A}\in K$ then $F(\mathfrak{A}) \in K'$ so
\[
	(A,(R^{\mathfrak{A}})_{R\in\tau},(\neg R^{\mathfrak{A}})_{R\in\tau},
	(x\ne y)^{\mathfrak{A}}) \in K'
\]
from which it follows that $\mathfrak{A} \vDash \psi$.
Thus, $K$ is definable in $\mathcal{L}(\mathcal{Q}_{\hom})$ so by Lemma
\ref{prel_lemma} we have
$\mathcal{L}(\mathcal{Q}_{\emb}) \leq \mathcal{L}(\mathcal{Q}_{\hom})$,
and since every homomorphism-closed quantifier is also embedding-closed we have
$\mathcal{L}(\mathcal{Q}_{\emb}) \equiv \mathcal{L}(\mathcal{Q}_{\hom})$.
\end{proof}

@@

"$\Rightarrow$" Suppose we have
$\Psi(\mathfrak{A}_i), \overline{a} \vDash \bigvee T'$ for some $i \geq k'$
and $\overline{a} \in A_i^m$.
Then for some $j$ with $k \leq j \leq k'$ there is $\overline{a} \in A_j^m$ such that
$\mathfrak{A}_j, \overline{a} \vDash \bigvee T$.
By (\ref{eq_0}) it follows that $\Psi(\mathfrak{A}_j), \overline{a} \vDash \varphi$
and, since $\Psi(\mathfrak{A}_j) \leq \Psi(\mathfrak{A}_i)$, we have
$\Psi(\mathfrak{A}_i), \overline{a} \vDash \varphi$ and thus
$\mathfrak{A}_i, \overline{a} \vDash \bigvee T$.

"$\Leftarrow$" Let $\mathfrak{A}_i, \overline{a} \vDash \bigvee T$
for some $i \geq k'$ and $\overline{a} \in A_i^m$.
Assume to the contrary that
$\Psi(\mathfrak{A}_i), \overline{a} \nvDash \bigvee T'$.
Then the atomic type $t$ of $\sigma$ for which
$\Psi(\mathfrak{A}_i), \overline{a} \vDash t$ is in $T_i$ but not in $T_k'$
which contradicts the set's $T_k'$ maximality.
The claim (\ref{eq_1}) thus holds.

>>>>

\begin{remark}
In addition to being closed under embeddings, we can imagine vari\-ous other
interesting closure conditions for quantifiers, for example, being closed under
homomorphisms. 
The rest of this article will be restricted to the study of 
embedding-closed quantifiers only.
However, we would like to mention here that it can be shown 
that any embedding-closed property can
be defined in a logic with only homomorphism-closed quantifiers.
Thus, requiring quantifiers to be closed under homomorphisms is not essentially
more restrictive than the requirement that they are closed under embeddings.
\end{remark}

@@

First, suppose we have
$\Psi(\mathfrak{A}_i), \overline{a} \vDash \bigvee T'$ for some $i \geq k'$
and $\overline{a} \in A_i^m$.
Then for some $j$ with $k \leq j \leq k'$ there is $\overline{a} \in A_j^m$ 
such that $\mathfrak{A}_j, \overline{a} \vDash \bigvee T$.
By (\ref{eq_0}) it follows that 
$\Psi(\mathfrak{A}_j), \overline{a} \vDash \varphi$
and, since $\Psi(\mathfrak{A}_j) \leq \Psi(\mathfrak{A}_i)$, we have
$\Psi(\mathfrak{A}_i), \overline{a} \vDash \varphi$ and thus
$\mathfrak{A}_i, \overline{a} \vDash \bigvee T$.

For the other direction, let $\mathfrak{A}_i, \overline{a} \vDash \bigvee T$
for some $i \geq k'$ and $\overline{a} \in A_i^m$, and
assume to the contrary that
$\Psi(\mathfrak{A}_i), \overline{a} \nvDash \bigvee T'$.
Then the atomic type $t$ of $\sigma$ for which
$\Psi(\mathfrak{A}_i), \overline{a} \vDash t$ is in $T_i$ but not in $T_k'$
which contradicts the set's $T_k'$ maximality.
The claim (\ref{eq_1}) thus holds.
